Six ways to understand monads\\ \small{Experiences from the functional programming course at TU Delft}

### 2012-10-30 latex/paper.tex
More content + refs

\author{\IEEEauthorblockN{Erik Meijer}
\and
\IEEEauthorblockN{Georgios Gousios}
\IEEEauthorblockA{
Delft University of Technology\\
Delft, The Netherlands\\
Email: {H.J.M.Meijer, G.Gousios, Arie.vanDeursen@tudelft.nl} 
}  
\and
\IEEEauthorblockN{Arie van Deursen}
}
\maketitle

@@

The aim of the course is to teach the principles of pure functional
programming, and the corresponding Category theoretical principles, using the
Haskell programming language. More specifically, the educational purposes
of the course were:

@@

@@

The course consisted of a series of lectures, of which two where invited
lectures by an external instructor, laboratories, a series of intermediate
assignments and a final assignment. The students would be evaluated by their
performance on the final assignment only.

@@

\cite{Hutto07}

@@

\subsection{The labs}

@@

\subsection{Teaching Monads}

@@

>>>>

\author{\IEEEauthorblockN{Erik Meijer}
\and
\IEEEauthorblockN{Georgios Gousios}
\IEEEauthorblockA{
Delft University of Technology\\
Delft, The Netherlands\\
Email: {H.J.M.Meijer, G.Gousios@tudelft.nl} 
}  
}
\maketitle

@@

The aim of the course was to teach the principles of pure functional
programming, and the corresponding Category theoretical principles. 
More specifically, the educational purposes of the course were:

@@

  \item To

@@

The course consisted of a series of lectures, of which two where invited
lectures by an external instructor, and laboratory sessions. In total, the
course consisted of 14 hours of lectures and 8 hours of lab work.
The students
also had to carry out a series of intermediate assignments and a final project.
The students would be evaluated by their performance on the final assignment. 
The lectures took place in a quarter (half semester), during 3 intense 
weeks; the students had another 6 weeks to finish their final projects.

@@

The lectures covered the following topics:

  \item Functions and functional composition, higher order functions, recursive
    functions, avoiding recursion

  \item Basic types, function types, currying

  \item Lists, mapping and traversal using folds

  \item Monads, composition and their application on the {\sc linq} query 
    language~\cite{Meije11}

  \item Monoids, functors and their application on data structure processing

  \item Functional formalization of the Map/Reduce data processing
    paradigm~\cite{Lamme08}

The lecture material was loosely based on Graham Hutton's functional programming
course at the University of Nottingham. The course's recommended book was
Programming in Haskell by the same author~\cite{Hutto07}.

@@

\subsection{Labs}

The purpose of the lab sessions where to help students carry out their final
assignments and explain better concepts that might not be clear to students. 
Students convened every week, presented their progress and discussed with 
each other and with the lab demonstrator design and practical issues that
arised in the implementation of their projects. To further motivate students
to work on their projects, one of the lab sessions was converted to a full
day coding sprint.

@@

\subsection{Teaching by example}

@@

\subsection{The coding sprint day}

Coding sprints or ``hackathons'' have long been used by open source software
projects to speed up development and increase participation of community
members~\cite{Adams11}. In a typical sprint, project members set measurable
targets and work intensively to
